---
title: "Untitled"
output: html_document
---

```{r}
library(readr)
CDM_IM_country_level_v0_9 <- read_csv("~/ImmigrationDUR/Data/CDM-IM_country-level_v0.9.csv")
#View(CDM_IM_country_level_v0_9)

library(readr)
usa <- read_csv("~/ImmigrationDUR/usa_00002.csv")

library(dplyr)
library(plyr)

usa1 = usa['BPLD' == 100]

plyr::count(usa, 'MIGRATE1D')
```


Cleaning
```{r}
library(readr)
library(mosaic)
library(dplyr)
library(plyr)
usa <- read_csv("~/ImmigrationDUR/Josh_files/usa_00003.csv")
#View(usa_00003)

usa1 <- usa[usa$BPLD == 60012 | usa$BPLD == 52200 | usa$BPLD == 60013 | usa$BPLD == 54400 | usa$BPLD == 60015 | usa$BPLD == 53200 | usa$BPLD == 60028 | usa$BPLD == 50210 | usa$BPLD == 60053 | usa$BPLD == 54100,]

usa1 <-usa1[usa$YRIMMIG >2004,] #REMOVES 41760 datapoints
usa1 <- usa1 %>% mutate(Birthplace = ifelse(BPLD == 60012,"Egypt",
                                    ifelse(BPLD == 52200, "Iran",
                                    ifelse(BPLD == 60013, "Libya",
                                    ifelse(BPLD == 54400, "Yemen",
                                    ifelse(BPLD == 60015, "Sudan",
                                    ifelse(BPLD == 53200, "Iraq",
                                    ifelse(BPLD == 60028, "Mali",
                                    ifelse(BPLD == 50210, "North Korea",                                                         ifelse(BPLD == 60053, "Somalia",
                                    ifelse(BPLD == 54100, "Syria",NA)))))))))))


usa1 <- usa1 %>% mutate(Education = ifelse(GRADEATT == 0,"NA",
                                    ifelse(GRADEATT == 1, "Nursury/Preschool",
                                    ifelse(GRADEATT == 2, "Kindergarten",
                                    ifelse(GRADEATT == 3, "Grade 1-4",
                                    ifelse(GRADEATT == 4, "Grade 5-8",
                                    ifelse(GRADEATT == 5, "Grade 9-12",
                                    ifelse(GRADEATT == 6, "College Undergrad",
                                    ifelse(GRADEATT == 7, "Graduate or Professional",NA)))))))))

usa1 <- usa1 %>% mutate(Migrate = ifelse(MIGRATE1 == 0,"NA",
                                    ifelse(MIGRATE1 == 1, "Same House",
                                    ifelse(MIGRATE1 == 2, "Moved Within State",
                                    ifelse(MIGRATE1 == 3, "Moved Between States",
                                    ifelse(MIGRATE1 == 4, "Abroad 1 Year",
                                    ifelse(MIGRATE1 == 9, "Unknown" , "Other")))))))



```


Education
```{r}
usa_education <- usa1[usa1$Education != "NA",] #REMOVES 11640 Observations, 1625834 Individuals

#usa_education <- usa1


expanded <- data.frame(Education = rep(usa_education$Education, usa_education$PERWT),
                       Birthplace = rep(usa_education$Birthplace, usa_education$PERWT),
                       GRADEATT = rep(usa_education$GRADEATT, usa_education$PERWT))

plyr::count(expanded, "Education")

library(ggplot2)
library(plotly)
#ZIP??
ggplot(expanded, aes(GRADEATT)) + geom_bar()

g <- ggplot(expanded, aes(GRADEATT, fill= Birthplace)) + geom_bar()

ggplotly(g)
plyr::count(expanded, "Birthplace")

plyr::count(expanded, "Education")
ctable <- table(expanded[,1:2])
ctable


```

Education cont.
```{r}
g_density <- ggplot(expanded, aes(GRADEATT, fill = Birthplace, colour = Birthplace)) + geom_density(alpha=.5)
ggplotly(g_density)

aggregate(expanded[, 3], list(expanded$Birthplace), mean)
aggregate(expanded[, 3], list(expanded$Birthplace), median)

g_box <- ggplot(expanded, aes(x=Birthplace, y=GRADEATT)) + geom_boxplot()
ggplotly(g_box)
```
Need to look at age as well.


Migrate
```{r}
usa_migrate <- usa1[usa1$Migrate != "NA",] #REMOVES 57 Observations, 7348 Individuals

#usa_migrate <- usa1


migrate_expanded <- data.frame(Migrate = rep(usa_migrate$Migrate, usa_migrate$PERWT),
                       Birthplace = rep(usa_migrate$Birthplace, usa_migrate$PERWT),
                       MIGRATE1 = rep(usa_migrate$MIGRATE1, usa_migrate$PERWT))

plyr::count(migrate_expanded, "Migrate")

library(plotly)

#ZIP BUT WITH 1?
ggplot(migrate_expanded, aes(MIGRATE1)) + geom_bar()

g <- ggplot(migrate_expanded, aes(MIGRATE1, fill= Birthplace)) + geom_bar()

ggplotly(g)

plyr::count(migrate_expanded, "Birthplace")
plyr::count(migrate_expanded, "Migrate")

ctable <- table(migrate_expanded[,1:2])
ctable


```

Migrate cont.
```{r}
g_density <- ggplot(migrate_expanded, aes(MIGRATE1, fill = Birthplace, colour = Birthplace)) + geom_density(alpha=.5)
ggplotly(g_density)

aggregate(migrate_expanded[, 3], list(migrate_expanded$Birthplace), mean)
aggregate(migrate_expanded[, 3], list(migrate_expanded$Birthplace), median)

g_box <- ggplot(migrate_expanded, aes(x=Birthplace, y=MIGRATE1)) + geom_boxplot()
g_box
```

One reason that it does not appear that there is a lot of migration, is because migrants may not have come in at the same time, and many might have come in later (after 2011). Migrants may not have had "time" to move to a different area. Explore the years next. 

Maybe use # of years a person is in the US


```{r}
usa_year <- usa1[!is.na(usa1$YRIMMIG),] #REMOVES 41760 Observations, 1625834 Individuals

#usa_year <- usa1


year_expanded <- data.frame(Year = rep(usa_year$YRIMMIG, usa_year$PERWT),
                       Birthplace = rep(usa_year$Birthplace, usa_year$PERWT))

plyr::count(year_expanded, "Year")

library(ggplot2)
library(plotly)
#ZIP??
ggplot(year_expanded, aes(Year)) + geom_bar()

g <- ggplot(year_expanded, aes(Year, fill= Birthplace)) + geom_bar()

ggplotly(g)
plyr::count(year_expanded, "Birthplace")

plyr::count(year_expanded, "Year")
ctable <- table(year_expanded[,1:2])
ctable


g_density <- ggplot(year_expanded, aes(Year, fill = Birthplace, colour = Birthplace)) + geom_density(alpha=.5)
g_density

aggregate(year_expanded[, 1], list(year_expanded$Birthplace), mean)
aggregate(year_expanded[, 1], list(year_expanded$Birthplace), median)

g_box <- ggplot(year_expanded, aes(x=Birthplace, y=Year)) + geom_boxplot()
g_box
```


WE SHOULDN'T USE CPS AND WE WILL US IPUMS-USA
```{r}
library(readr)
cps_00003 <- read_csv("~/ImmigrationDUR/Josh_files/cps_00003.csv")
cps_00003 <-cps_00003[cps_00003$YRIMMIG >2004,] #REMOVES 41760 datapoints

cps <- cps_00003

cps <- cps[cps$BPL == 60012 | cps$BPL == 53000 | cps$BPL == 60019 | cps$BPL == 54400 | cps$BPL == 60018 | cps$BPL == 53200 | cps$BPL == 60053 | cps$BPL == 54100,]


cps <- cps %>% mutate(Birthplace = ifelse(BPL == 60012,"Egypt",
                                    ifelse(BPL == 53000, "Iran",
                                    ifelse(BPL == 60019, "Libya",
                                    ifelse(BPL == 54400, "Yemen",
                                    ifelse(BPL == 60018, "Sudan",
                                    ifelse(BPL == 53200, "Iraq",
                                    ifelse(BPL == 60053, "Somalia",
                                    ifelse(BPL == 54100, "Syria",NA)))))))))



plyr::count(cps, "Birthplace")

ggplot(cps, aes(MIGSTA1)) + geom_bar()

plyr::count(cps, "MIGSTA1")
```


https://www.migrationpolicy.org/programs/data-hub/us-immigration-trends#source


