---
title: "American Dream"
author: "Pony Ameri"
date: "3/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Cleaning
```{r}
library(readr)
library(dplyr)
library(plyr)
library(car)
library(ipumsr)
library(labelled)
library(ggplot2)
library(mosaic)



usa_ddi <- read_ipums_ddi("/home/pelayo1/ImmigrationDUR/Josh_files/usa_00008.xml")
usa_data <- read_ipums_micro(usa_ddi, verbose = FALSE)
usa_data <- usa_data %>%
  mutate(METRO = droplevels(as_factor(METRO)), BIRTHPLACE_factor= droplevels(as_factor(BPL)), LABFORCE1= droplevels(as_factor(LABFORCE)), SEX= droplevels(as_factor(SEX)),FOODSTMP= droplevels(as_factor(FOODSTMP)), RELATE= droplevels(as_factor(RELATE)),RELATED= droplevels(as_factor(RELATED)), MARST= droplevels(as_factor(MARST)), BPL = droplevels(as_factor(BPL)),BPLD = droplevels(as_factor(BPLD)), CITIZEN = droplevels(as_factor(CITIZEN)),RACESING=droplevels.factor(RACESING),RACESINGD=droplevels(as_factor(RACESINGD)),HCOVANY = droplevels(as_factor(HCOVANY)),HCOPRIV = droplevels(as_factor(HCOVPRIV)), EDUCATION = droplevels(as_factor(GRADEATT)),EDUCATIOND= droplevels(as_factor(GRADEATTD)),EMPLOYMENT = droplevels(as_factor(EMPSTAT)), EMPLOYMENTD = droplevels(as_factor(EMPSTATD)), LABFORCE = droplevels(as_factor(LABFORCE)), IND = droplevels(as_factor(IND)),
 CLASSWKR = droplevels(as_factor(CLASSWKR)), CLASSWKRD = droplevels(as_factor(CLASSWKRD)), POVERTY = droplevels(as_factor(POVERTY)), MIGRATE1 = droplevels(as_factor(MIGRATE1)), MIGRATE1D = droplevels(as_factor(MIGRATE1D)), MIGPLAC1 = droplevels(as_factor(MIGPLAC1)), TRANWORK = droplevels(as_factor(TRANWORK)))
usa_data$BPLD <-recode_factor(usa_data$BPLD, 'Egypt/United Arab Rep.' = "Egypt/UAE", 'Yemen Arab Republic (North)' = "Yemen")
usa_data$EMPLOYMENT <-recode_factor(usa_data$EMPLOYMENT, 'N/A' = "Armed Forces")
usa_data$LABFORCE1<-recode_factor(usa_data$LABFORCE1,'Yes, in the labor force'= "Yes",'No, not in the labor force'="No") 
usa_data$METRO<-recode_factor(usa_data$METRO,'Central / Principal city status unknown' = " Metro area, central", 'In metro area, central / principal city' = " Metro area, central",'In metro area, outside central / principal city' ="Metro area, outside central")
  
usa_data <- usa_data %>% filter(YRIMMIG >1995,AGE >= 25)
usa_data$YRSINUSA <- usa_data$YEAR - usa_data$YRIMMIG
usa_data <- usa_data[usa_data$OWNERSHP != "0",]
usa_data$POVERTY <- as.numeric(usa_data$POVERTY)

usa_data <- usa_data %>% mutate(YRGROUP = ifelse(YRSINUSA <= 5,"0-5",
                                    ifelse(YRSINUSA >5 & YRSINUSA <=10, "6-10",
                                   ifelse(YRSINUSA >10 & YRSINUSA <=15, "11-15",
                                    ifelse(YRSINUSA >15, ">15", "N/A")))),
                                Employment = ifelse(EMPLOYMENT == "Employed", 1,0), 
                              Laborforce = ifelse(LABFORCE1=="Yes",1,0),
                              Residence = ifelse(METRO == "Metro area, outside  central",1,0), 
                              Ownhome = ifelse(OWNERSHP==1, 1,0),  
                              OWNERSHP = droplevels(as_factor(OWNERSHP)),
                              Poverty = ifelse(POVERTY >100, 1,0))
usa_data$OWNERSHP <-recode_factor(usa_data$OWNERSHP, 'Owned or being bought (loan)' = "Own", 'Rented' = "Rent")
View(usa_data)
table(usa_data$HCOVANY)

```

```{r}
usa_education <- usa_data[usa_data$EDUCATIOND != "N/A",] #REMOVES 11640 Observations, 1625834 Individuals

#usa_education <- usa1


expanded <- data.frame(Education = rep(usa_education$EDUCATIOND, usa_education$PERWT),
                       Birthplace = rep(usa_education$BPLD, usa_education$PERWT),
                       GRADEATT = rep(usa_education$GRADEATT, usa_education$PERWT))

plyr::count(expanded, "Education")

#ZIP??
ggplot(expanded, aes(GRADEATT)) + geom_bar()

g <- ggplot(expanded, aes(GRADEATT, fill= Birthplace)) + geom_bar()
g
plyr::count(expanded, "Birthplace")

plyr::count(expanded, "Education")
ctable <- table(expanded[,1:2])
ctable

g_density <- ggplot(expanded, aes(GRADEATT, fill = Birthplace, colour = Birthplace)) + geom_density(alpha=.5)
g_density

aggregate(expanded[, 3], list(expanded$Birthplace), mean)
aggregate(expanded[, 3], list(expanded$Birthplace), median)

g_box <- ggplot(expanded, aes(x=Birthplace, y=GRADEATT)) + geom_boxplot()
g_box
```
This is for age >25. Education is not a great predictor. 

Migrate
```{r}
usa1 <- usa_data
usa_migrate <- usa1[usa1$Migrate != "N/A",] #REMOVES 57 Observations, 7348 Individuals

#usa_migrate <- usa1


migrate_expanded <- data.frame(Migrate = rep(usa_migrate$MIGRATE1, usa_migrate$PERWT),
                       Birthplace = rep(usa_migrate$Birthplace, usa_migrate$PERWT),
                       MIGRATE1 = rep(usa_migrate$MIGRATE1, usa_migrate$PERWT))

plyr::count(migrate_expanded, "Migrate")

library(plotly)

#ZIP BUT WITH 1?
ggplot(migrate_expanded, aes(MIGRATE1)) + geom_bar()

g <- ggplot(migrate_expanded, aes(MIGRATE1, fill= Birthplace)) + geom_bar()

ggplotly(g)

plyr::count(migrate_expanded, "Birthplace")
plyr::count(migrate_expanded, "Migrate")

ctable <- table(migrate_expanded[,1:2])
ctable


```

Migrate cont.
```{r}
g_density <- ggplot(migrate_expanded, aes(MIGRATE1, fill = Birthplace, colour = Birthplace)) + geom_density(alpha=.5)
ggplotly(g_density)

aggregate(migrate_expanded[, 3], list(migrate_expanded$Birthplace), mean)
aggregate(migrate_expanded[, 3], list(migrate_expanded$Birthplace), median)

g_box <- ggplot(migrate_expanded, aes(x=Birthplace, y=MIGRATE1)) + geom_boxplot()
g_box
```

One reason that it does not appear that there is a lot of migration, is because migrants may not have come in at the same time, and many might have come in later (after 2011). Migrants may not have had "time" to move to a different area. Explore the years next. 

Maybe use # of years a person is in the US


```{r}
usa1 <- usa_data
usa_year <- usa1[!is.na(usa1$YRIMMIG),] #REMOVES 41760 Observations, 1625834 Individuals

#usa_year <- usa1


year_expanded <- data.frame(Year = rep(usa_year$YRIMMIG, usa_year$PERWT),
                       Birthplace = rep(usa_year$BPLD, usa_year$PERWT))

plyr::count(year_expanded, "Year")

library(ggplot2)
library(plotly)
#ZIP??
ggplot(year_expanded, aes(Year)) + geom_bar()

g <- ggplot(year_expanded, aes(Year, fill= Birthplace)) + geom_bar()

ggplotly(g)
plyr::count(year_expanded, "Birthplace")
plyr::count(year_expanded, "Year")
ctable <- table(year_expanded[,1:2])
ctable


g_density <- ggplot(year_expanded, aes(Year, fill = Birthplace, colour = Birthplace)) + geom_density(alpha=.5)
g_density
ggplotly(g_density)

aggregate(year_expanded[, 1], list(year_expanded$Birthplace), mean)
aggregate(year_expanded[, 1], list(year_expanded$Birthplace), median)

g_box <- ggplot(year_expanded, aes(x=Birthplace, y=Year)) + geom_boxplot()
g_box
```


```{r, employment status}
#usa_year <- usa1
usa1<-usa_data

employ_expanded <- data.frame(Country = rep(usa1$BPLD, usa1$PERWT),
                       Employment = rep(usa1$EMPLOYMENT, usa1$PERWT),Year=rep(usa1$YEAR, usa1$PERWT), Gender=rep(usa1$SEX, usa1$PERWT), Time = rep(usa1$YRSINUSA, usa1$PERWT))
ggplot(employ_expanded, aes(Country, fill= Employment)) + geom_bar()
ggplot(employ_expanded, aes(Country, fill= Employment)) + geom_density(alpha=.5)
table(employ_expanded[,1:2])


ggplot(employ_expanded, aes(Time, fill=Employment)) + 
  geom_bar() +labs(title="Yearly Time Series",subtitle ="Employment over time", x = "Year", y ="Count")  + theme(plot.title = element_text(lineheight=.8, face="bold", 
                                          size = 15)) +
           theme(text = element_text(size=12))

ggplot(employ_expanded, aes(Country, fill=Employment)) + 
  geom_bar() +facet_wrap(~Time)+ labs(title="Yearly Time Series",subtitle ="Employment over time", x = "Year", y ="Count")  + theme(plot.title = element_text(lineheight=.8, face="bold", 
                                          size = 15)) +
           theme(text = element_text(size=12))

```

```{r, labor force}
#usa_year <- usa1
usa1<-usa_data
lbf_expanded <- data.frame(Country = rep(usa1$BPLD, usa1$PERWT),
                       Laborforce = rep(usa1$LABFORCE1, usa1$PERWT), Year=rep(usa1$YEAR, usa1$PERWT), Time = rep(usa1$YRSINUSA, usa1$PERWT) )
ggplot(lbf_expanded, aes(Country, fill= Laborforce)) + geom_bar()

ggplot(lbf_expanded, aes(Time, fill= Laborforce)) + 
  geom_bar(position = position_dodge(0.9), width = 0.9) +labs(title= "Labor Force Status over Time", x = "Year", y ="Count")  + theme(plot.title = element_text(lineheight=.8, face="bold", 
                                          size = 15)) +
           theme(text = element_text(size=12))
```

```{r}
usa1<-usa_data
CW_expanded <- data.frame(Country = rep(usa1$BPLD, usa1$PERWT),
                       Classwork = rep(usa1$CLASSWKR, usa1$PERWT), Year=rep(usa1$YEAR, usa1$PERWT), Time = rep(usa1$YRSINUSA, usa1$PERWT) )
ggplot(CW_expanded, aes(Country, fill= Classwork)) + geom_bar()
ggplot(CW_expanded, aes(Classwork, fill= Country)) + geom_bar()
ggplot(CW_expanded, aes(Country, fill= Classwork)) + geom_density(alpha=.5)

ggplot(CW_expanded, aes(Time, fill= Classwork)) + 
  geom_bar() +labs(title= "Class of worker over Time", x = "Year", y ="Count")  + theme(plot.title = element_text(lineheight=.8, face="bold", 
                                          size = 15)) +
           theme(text = element_text(size=12))
```

ggplot(home, aes(Year,fill =City)) + geom_bar ()+facet_wrap(~Country)
```{r}
usa1<-usa_data
home <- data.frame(Country = rep(usa1$BPLD, usa1$PERWT),
                       Area = rep(usa1$METRO, usa1$PERWT), Year=rep(usa1$YEAR, usa1$PERWT), Time= rep(usa1$YRSINUSA, usa1$PERWT))
ggplot(home, aes(Country, fill= Area)) + geom_bar(position = position_dodge(0.8))
ggplot(home, aes(Country, fill= Area)) + geom_density(alpha=.5)
plyr::count(home, "Area")

ggplot(home, aes(Time, fill= Area)) + 
  geom_bar(position = position_dodge(0.9), width = 0.9) +labs(title= "Area of Residence over time", x = "Year", y ="Count")  + theme(plot.title = element_text(lineheight=.8, face="bold", 
                                          size = 15)) +
           theme(text = element_text(size=12))
```

```{r, Income}
usa1<-usa_data

cps_data %>% 
  filter(FOODSTMP %in% c(1,2)) %>% 
  group_by((EMPSTAT)) %>% 
  summarize(fs_pct = weighted.mean(FOODSTMP == 2, ASECWT))
  
Income <- data.frame(Country = rep(usa1$BPLD, usa1$PERWT),
                       income = rep(usa1$FTOTINC, usa1$PERWT), Year=rep(usa1$YEAR, usa1$PERWT),Time= rep(usa1$YRSINUSA, usa1$PERWT))

Income3 <- aggregate( Income[,2], list(country=Income$Country, time= Income$Time,year =Income$Year),mean)
Income3$meanincome <- Income3$x

head(Income3)
ggplot(Income3, aes(x=year,y=meanincome)) + 
  geom_line(aes(color = country), size = 1) +labs(title="Yearly Time Series",subtitle ="Income over time by country", x = "Year", y ="Income")  + theme(plot.title = element_text(lineheight=.8, face="bold", 
                                          size = 15)) +
           theme(text = element_text(size=12))

# Area plot
ggplot(Income3, aes(time,y=meanincome)) + 
  geom_area(aes(color = country, fill = country), 
            alpha = 0.5, position = position_dodge(0.8))+labs(title="Yearly Time Series",subtitle ="Income over time by country", x = "Year", y ="Income")  + theme(plot.title = element_text(lineheight=.8, face="bold", 
                                          size = 15)) +
           theme(text = element_text(size=12))
```

Health Care

```{r}

usa1 <- usa_data
usa_health <- usa1[!is.na(usa1$HCOVANY),] #REMOVES 842 Observations, 1625834 Individuals

#usa_education <- usa1


health_expanded <- data.frame(Health = rep(usa_health$HCOVANY, usa_health$PERWT),
                       Birthplace = rep(usa_health$BPLD, usa_health$PERWT),
                       Private = rep(usa_health$HCOPRIV, usa_health$PERWT))

plyr::count(health_expanded, "Health") #4431 without insurance, 12693-4941 = 7752 with public health insurance
plyr::count(health_expanded, "Private") #17124 total; 4941 with private health insurance



#ZIP??
ggplot(health_expanded, aes(Health)) + geom_bar()

g <- ggplot(health_expanded, aes(Health, fill= Private)) + geom_bar()
g

ctable <- table(health_expanded[,1:2])

ctable

require(ggplot2)
require(scales)

ggplot(health_expanded, aes(x= Birthplace,  group=Health)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
    geom_text(aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.5) +
    labs(y = "Percent", fill="day") +
    facet_grid(~Health) +
    scale_y_continuous(labels = scales::percent)

aggregate(expanded[, 3], list(expanded$Birthplace), mean)
aggregate(expanded[, 3], list(expanded$Birthplace), median)
```

```{r}


```

