---
title: "Analysis"
author: "So Mang Han"
date: "4/5/2018"
output: html_document
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE)
```


```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(readxl)
library(randomForest)
library(FNN)
library(stats)
library(ggfortify)
library(party)
library(mosaic)
install.packages("factoextra")
suppressMessages(library(MASS))
suppressMessages(library(factoextra))

```


##EDA for the first question 
```{r}
Immigrant_2005 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2005.csv")
Immigrant_2006 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2006.csv")
Immigrant_2007 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2007.csv")
Immigrant_2008 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2008.csv")
Immigrant_2009 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2009.csv")
Immigrant_2010 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2010.csv")
Immigrant_2011 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2011.csv")
Immigrant_2012 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2012.csv")
Immigrant_2013 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2013.csv")
Immigrant_2014 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2014.csv")
Immigrant_2015 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2015.csv")
Immigrant_2016 <- read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/Immigrant_2016.csv")

all_immigrant_year=rbind(Immigrant_2005,Immigrant_2006,Immigrant_2007,Immigrant_2008,Immigrant_2009,Immigrant_2010,Immigrant_2011,Immigrant_2012,Immigrant_2013,Immigrant_2014,Immigrant_2015,Immigrant_2016)

all_immigrant_year=as.data.frame(all_immigrant_year)
all_immigrant_year$Immigrant=as.numeric(all_immigrant_year$Immigrant)

mean_immigrant=all_immigrant_year %>% group_by(State) %>% summarise(mean_number=mean(Immigrant,na.rm=T))

states <- map_data("state")
political_affliation_states=data.frame(state=c(unique(states$region)[1],"alaska",unique(states$region)[2:10],"hawaii",unique(states$region)[11:49]),color=c(rep(c("red"),4),rep(c("blue"),5),rep(c("red"),2),rep(c("blue"),1),rep(c("red"),1),rep(c("blue"),1),rep(c("red"),5),rep(c("blue"),3),rep(c("red"),1),rep(c("blue"),1),rep(c("red"),4),rep(c("blue"),5),rep(c("red"),4),rep(c("blue"),1),rep(c("red"),1),rep(c("blue"),1),rep(c("red"),5),rep(c("blue"),3),rep(c("red"),3)))

mean_immigrant$State=tolower(mean_immigrant$State)

political_affliation_states$state=as.character(political_affliation_states$state)

mean_numbers=data.frame(mean_number_immigrant=mean_immigrant[which(is.element(mean_immigrant$State,political_affliation_states$state)==T),])


names(mean_numbers)=c("state","mean_number")

state_data=cbind(mean_numbers,political_affliation_states$color)

Christan_percent=c(86,62,67,79,63,64,70,69,65,70,79,63,67,71,72,77,76,76,84,60,69,58,70,74,83,77,65,75,66,59,67,75,60,77,77,73,79,61,73,75,78,79,81,77,73,54,73,61,78,71,71)

demography_us<-read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/raw_data.csv")
demography_us_new=data.frame(state=demography_us$Location[2:52],white_population=demography_us$White[2:52],christan_percent=Christan_percent)
demography_us_new=mutate(demography_us_new,ifelse(demography_us_new$white_population>=0.7,"predominately_white","not_predominately_white"))
demography_us_new=mutate(demography_us_new,ifelse(demography_us_new$christan_percent>=mean(demography_us_new$christan_percent),"predominately_christian","not_predominately_christian"))

names(demography_us_new)[c(4,5)]=c("white_population","christian_population")

state_data_final=cbind(state_data,data1[,c(2,3)])
state_data_final=as.data.frame(state_data_final)


names(state_data_final)[3]=c("political_affliation")

state_data_final##the final data(for categorical analysis) 

demography_us_new##the final(for quantitative analysis)

```


```{r}
rownames(demography_us_new)=demography_us_new$state
demography_us_new_1=demography_us_new[,-1]


```


```{r}
##political ideology 
#liberal$mean_number_immigrant.mean_immigrant_number
liberal=filter(state_data_final,political_affliation=="blue")
conservative=filter(state_data_final,political_affliation=="red")
favstats(liberal$mean_number);favstats(conservative$mean_number)##comparing the mean number of immigrants by red states vs. blue states

ggplot(state_data_final, aes(x=political_affliation, y=mean_number,fill=political_affliation)) + 
  geom_boxplot()+labs(title="The mean of the number of immigrants from travel ban 8 countries during 2005-2016",x="political affliation", y = "the mean of immigrants accepted")+scale_fill_manual(values=c("blue", "red")) 

```
The mean of the number of immigrants accepted by blue states is higher than the mean of the number of immigrants accepted by red states.


```{r}
pre_dominately_white=filter(state_data_final,white_population=="predominately_white")
not_pre_dominately_white=filter(state_data_final,white_population=="not_predominately_white")
boxplot(pre_dominately_white$mean_number,not_pre_dominately_white$mean_number)

favstats(pre_dominately_white$mean_number);favstats(not_pre_dominately_white$mean_number)##comparing the mean number of immigrants by pre_dominately_white states vs. not_pre_dominately_white states

ggplot(state_data_final, aes(x=white_population, y=mean_number,fill=white_population)) + 
  geom_boxplot()+labs(title="The Mean of the Number of Immigrants From Travel Ban 8 Countries During 2005-2016",x="White Population", y = "The Mean of Immigrants Accepted")+scale_fill_manual(values=c("darkgreen", "purple")) 


```

The mean of the number of immigrants accepted by not predominately white states is higher than the mean of the number of immigrants accepted by predominately white.



```{r}
##political affliation and chrisitan population 

p <- ggplot(data=state_data_final, aes(x=political_affliation, y=mean_number, fill=christian_population)) +
    geom_bar(stat="identity", position=position_dodge())+xlab("Political Affliation") + ylab("Immigrant Number") + 
    ggtitle("Immigrant Number by Political Affliation and Chrisitan community")+
    scale_fill_manual(values=c("#999999", "#E69F00")) 
p

```

From the graph, we can tell that there is a tendency that not predominately chrisitan community is willing to accept more immigrants for both red and blue states.


```{r}
##christian population and white population 

p1 <- ggplot(data=state_data_final, aes(x=white_population, y=mean_number, fill=christian_population)) +
    geom_bar(stat="identity", position=position_dodge())+xlab("white population") + ylab("Refugee Number") + 
    ggtitle("Immigrant Number by White Population and Chrisitan Population")+
    scale_fill_manual(values=c("darkgreen", "darkred")) 

p1


```

Again,from the graph, we can tell that there is a tendency that not predominately chrisitan community is willing to accept more immigrants for both predominately white and not predominately white states.


```{r}
##political affliation and white population 

p2 <- ggplot(data=state_data_final, aes(x=white_population, y=mean_number, fill=political_affliation)) +
    geom_bar(stat="identity", position=position_dodge())+xlab("white population") + ylab("immigrant Number") + 
    ggtitle("immigrant Number by Political Affliation and White Population")+
    scale_fill_manual(values=c("blue", "red")) 

p2
```

It seems like political affliation does not significantly impact on the mean of immigrant number of predominately white state and the mean of immigrant number of not predominately white state




###IPUMS
```{r}

set.seed(360)
usa1=NULL;usa=NULL
usa <- read_csv("/home/pelayo1/ImmigrationDUR/American Dream/usa_00008.csv")

usa1 <- usa[usa$BPLD == 60012 | usa$BPLD == 52200 | usa$BPLD == 60013 | usa$BPLD == 54400 | usa$BPLD == 60015 | usa$BPLD == 53200 | usa$BPLD == 60028 | usa$BPLD == 50210 | usa$BPLD == 60053 | usa$BPLD == 54100,]

usa1 <-usa1[usa$YRIMMIG >2004,] 
usa1 <- usa1 %>% mutate(Birthplace = ifelse(BPLD == 60012,"Egypt",
                                    ifelse(BPLD == 52200, "Iran",
                                    ifelse(BPLD == 60013, "Libya",
                                    ifelse(BPLD == 54400 & BPLD == 54500, "Yemen",
                                    ifelse(BPLD == 60015, "Sudan",
                                    ifelse(BPLD == 53200, "Iraq",                                               ifelse(BPLD == 60053, "Somalia",
                                    ifelse(BPLD == 54100, "Syria",NA)))))))))


usa1 <- usa1 %>% mutate(Education = ifelse(GRADEATT == 0,"NA",
                                    ifelse(GRADEATT == 1, "Nursury/Preschool",
                                    ifelse(GRADEATT == 2, "Kindergarten",
                                    ifelse(GRADEATT == 3, "Grade 1-4",
                                    ifelse(GRADEATT == 4, "Grade 5-8",
                                    ifelse(GRADEATT == 5, "Grade 9-12",
                                    ifelse(GRADEATT == 6, "College Undergrad",
                                    ifelse(GRADEATT == 7, "Graduate or Professional",NA)))))))))




state=as.character(c(state.name[1:8],"District of Columbia",state.name[9:length(state.name)]))
state_index=seq(1,56,by=1)
state_index=state_index[c(-3,-7,-14,-43,-52)]
state_name=data.frame(state_n=state,index=state_index)

lines=NULL
immigrant_table=NULL
i=1
while(i<=length(state_index)){
lines=which(usa1$MIGPLAC1==state_name[i,2])
stuff=usa1[lines,]
stuff$living_state=as.character(state_name[i,1])
immigrant_table=rbind(immigrant_table,stuff)
i=i+1
}

table_final=immigrant_table[,-c(2,3,4,5,7,8,11,13,14,15,18,19,20,21,22,26,28,29,30,31,33,34,35,36,37,40,41,38,43,44,45,46,47,48,50,51)]



colnames(table_final)=c("Census_Year","Metropolitan_Status","Ownership of dwelling","Food_Stamp_Recipiency","Person_Weight","Sex","Age","Citizenship_Status","Year_of_Immigration","Single_Race_Detailed_Version","Any_Health_Insurance_Coverage","Employment_Status_General_Version","Total_Family_Income","Poverty_Status","Independent_Living_Difficulty","Means_of_Transportation_to_Work","Birthplace_final","Education_final","State_Residency_final")

```


```{r}
colnames(table_final)

```



```{r}
##making variables
states=NULL;political_affliation_states=NULL
states <- map_data("state")

political_affliation_states=data.frame(state=c(unique(states$region)[1],"alaska",unique(states$region)[2:10],"hawaii",unique(states$region)[11:49]),color=c(rep(c("red"),4),rep(c("blue"),5),rep(c("red"),2),rep(c("blue"),1),rep(c("red"),1),rep(c("blue"),1),rep(c("red"),5),rep(c("blue"),3),rep(c("red"),1),rep(c("blue"),1),rep(c("red"),4),rep(c("blue"),5),rep(c("red"),4),rep(c("blue"),1),rep(c("red"),1),rep(c("blue"),1),rep(c("red"),5),rep(c("blue"),3),rep(c("red"),3)))


Christan_percent=c(86,62,67,79,63,64,70,69,65,70,79,63,67,71,72,77,76,76,84,60,69,58,70,74,83,77,65,75,66,59,67,75,60,77,77,73,79,61,73,75,78,79,81,77,73,54,73,61,78,71,71)

demography_us<-read_csv("/home/pelayo1/ImmigrationDUR/Data/immigrant/raw_data.csv")

demography_us_new=data.frame(state=demography_us$Location[2:52],white_population=demography_us$White[2:52],christan_percent=Christan_percent)

demography_us_new=mutate(demography_us_new,ifelse(demography_us_new$white_population>=0.7,"predominately_white","not_predominately_white"))

demography_us_new=mutate(demography_us_new,ifelse(demography_us_new$christan_percent>=mean(demography_us_new$christan_percent),"predominately_christian","not_predominately_christian"))

head(demography_us_new)

names(demography_us_new)[c(4,5)]=c("Dominance_White","Dominance_Christan")



table_final$white_population=0;table_final$christan_percent=0;table_final$Dominance_White=0;table_final$Dominance_Christan=0

i=1
while(i<=nrow(table_final)){
 a=demography_us_new[which(table_final[i,]$State_Residency_final==demography_us_new$state),]
table_final$white_population[i]=a$white_population
table_final$christan_percent[i]=a$christan_percent
table_final$Dominance_White[i]=a$Dominance_White
table_final$Dominance_Christan[i]=a$Dominance_Christan
i=i+1
}

character_vars <- lapply(table_final, class) == "character"
table_final[, character_vars] <- lapply(table_final[, character_vars], as.factor)
factor_vars <- lapply(table_final, class) == "factor"
table_final[, factor_vars] <- lapply(table_final[, factor_vars], as.numeric)

##1-->not predominately white 2--> predominately white
##1-->not predominately christian 2--> predominately christian


```



```{r}
ggplot(table_final,aes(Total_Personal_Income,Grade_Level_Attending_General_Version,color=as.factor(Dominance_Christan)))+geom_point(aes(size = Person_Weight))+xlab("Personal Income") + ylab("Education Level") + 
    ggtitle("Classification of people by Education Level and Personal Income")
```



```{r}

##Data cleaning 

immigrant_table.df<-sapply(immigrant_table,as.numeric) 
immigrant_table.df<-as.data.frame(immigrant_table.df)

names(immigrant_table.df)[54:57]=c("White_Percent","Dominance_White","Christan_Percent","Dominance_Christan")
immigrant_table.df$Christan_Percent=0.01*immigrant_table.df$Christan_Percent

#get rid of unnessary variables
immigrant_table_new.df=immigrant_table.df[,-c(2,3,4,5,7,9,13,17,19,20,24,28,29,31,33,35,40,41,42,43,56,54)]

immigrant_table_new.df$Year_of_Staying=immigrant_table.df$Census_Year-immigrant_table.df$Year_of_Immigration
immigrant_table_new.df=immigrant_table_new.df[,-c(1,12)]

immigrant_table_new.df=immigrant_table_new.df[,c(1:32,34,33)]


immigrant_table_new.df$Person_Weight
immigrant_table_new.df.expanded <- immigrant_table_new.df.expanded[rep(row.names(immigrant_table_new.df), immigrant_table_new.df$Person_Weight),]


immigrant_table_new.df.expanded =immigrant_table_new.df.expanded [complete.cases(immigrant_table_new.df.expanded ), ]#no NA


```


#Regression models

I am going to create models to predict Chirstan percent with other 32 independent variables.  

##Linear 
```{r}
mod <- immigrant_table_new.df.expanded%>%
    lm(Christan_Percent  ~.,data=.)

summary(mod)

```


```{r}
pvals <- summary(mod)$coefficients[,4]
sigPVals <- pvals<0.05

sigFields <- names(immigrant_table_new.df.expanded)[sigPVals]
sigFields##statistically significant values 

```


```{r}
## CrossValidate the model for a given data frame.
mseCV <- function(data.df,kfolds=5){
    folds <- sample(1:kfolds,sampleSize,rep=T)
    mse <- rep(0,kfolds)
    for(k in 1:kfolds){
        train.df <- data.df[folds !=k,]
        test.df <- data.df[folds==k,]
        mod <- lm(Christan_Percent~.,data=train.df)
        vals <- predict(mod,newdata=test.df)
        mse[k] <- with(test.df,mean((Christan_Percent-vals)^2,na.rm = T))
    }
    mean(mse,na.rm = T)
}
```

##boostrapped version 
```{r}

sampleSize<-nrow(immigrant_table_new.df.expanded)
mseBoot <- function(data.df,M=30){
    mse <- rep(0,M)
    for(m in 1:M){
        bootSamp <- sample(1:sampleSize,sampleSize,rep=T)
        outOfBag <- setdiff(1:sampleSize,bootSamp)
        train.df <- data.df[bootSamp,]
        test.df <-   data.df[outOfBag,]
        mod <- lm(Christan_Percent~.,data=train.df)
        vals <- predict(mod,newdata=test.df)
        mse[m] <- with(test.df,mean((Christan_Percent-vals)^2,na.rm = T))
    }
    mean(mse,na.rm = T)
}
```


```{r}
##testing codes 
numPreds <- length(names(immigrant_table_new.df.expanded))-1
predIDs <- c(1,4,5,6,7,8,9:15)
predIDs <- c(1)

data.df=immigrant_table_new.df.expanded[,c(predIDs,numPreds+1)]

mseCV(data.df)

mseBoot(data.df)

```




##Forward subset  
```{r}

minMSE <- c()
currPreds <- c()
numPreds <- length(names(immigrant_table_new.df.expanded))-1
allPreds <- 1:(numPreds)
availPreds <- setdiff(allPreds,currPreds)
sampleSize <- nrow(immigrant_table_new.df.expanded)
maxPreds <- numPreds


while(length(currPreds) < maxPreds){
    ##add predictor which decreases error rate (as determined by CV or
    ##Bootstrapping)
    ## The MSEs computed as we add each of the available predictors
    allMSE <- c()
    for(id in availPreds){
        data.df <- immigrant_table_new.df.expanded[,c(currPreds,id,numPreds+1)]
        print(c(currPreds,id,numPreds+1))
        mse <- mseCV(data.df,5) ##5-fold to speed up things
        ##mse <- mseBoot(data.df,30)
        allMSE <- c(allMSE,mse)
    }
    ##Find the min
    id <- which.min(allMSE)
    ##get the best predictor and MSW
    bestPred <- availPreds[id]
    bestMSE <- min(allMSE,na.rm = T)
    ##Add these into the collection
    currPreds <- c(currPreds,bestPred)
    minMSE <- c(minMSE,bestMSE)
    availPreds <- setdiff(allPreds,currPreds)
    ## Print stuff out for debugging and attention-grabbing
    print(c(bestPred,bestMSE))
}
```

```{r}
descr <- with(immigrant_table_new.df.expanded,colnames(immigrant_table_new.df.expanded)[currPreds])
diffMSE <- c(0,minMSE[-length(minMSE)]-minMSE[-1])
result.df <- data.frame(id=1:length(descr),descr,minMSE,diffMSE=-10*diffMSE)
result.df <- result.df%>%
    mutate(descr=factor(descr,levels=rev(descr)))
ggplot(result.df,aes(id,minMSE))+
    geom_point(color="blue")+
    geom_line(color="black",size=0.25)+
    scale_x_continuous("Predictor Number")+
    scale_y_continuous("MSE")+
    ggtitle("Subset Selection: MSE Reduction")
```

```{r}
ggplot(result.df,
       aes(descr,minMSE))+
    geom_bar(stat="identity",fill="blue")+
    coord_flip()+
    ggtitle("Subset Selection: MSE Decrease")
```

```{r}
ggplot(result.df)+
    geom_segment(aes(x=id,xend=id,y=0,yend=minMSE),color="blue")+
    geom_point(aes(x=id,y=minMSE),color="blue",size=2)+
    scale_x_continuous(breaks=1:length(descr),label=descr)+
    coord_flip()+
    ggtitle("Subset Selection: MSE Decrease")
```

```{r}
ggplot(result.df,aes(descr,diffMSE))+
    geom_bar(stat="identity",fill="red")+
    coord_flip()+
    ggtitle("Subset Selection: MSE Chage")
```

```{r}
ggplot(result.df)+
    geom_segment(aes(x=id,xend=id,y=0,yend=diffMSE),color="red")+
    geom_point(aes(x=id,y=diffMSE),color="red",size=2)+
    scale_x_continuous(breaks=1:length(descr),label=descr)+
    coord_flip()+
    ggtitle("Subset Selection: MSE Decrease")
```




```{r}


head(immigrant_table_new.df.expanded)
summarised<-ggplot(immigrant_table_new.df.expanded, aes(x=sex)) +geom_bar(aes(fill=factor(sex)))+
    facet_wrap(age_cat~race)
summarised
```


```{r}
#get rid of unnessary variables
immigrant_table_new.df_=immigrant_table.df[,-c(2,3,4,5,7,9,13,17,19,20,24,28,29,31,33,35,40,41,42,43,56,54)]

immigrant_table_new.df$Year_of_Staying=immigrant_table.df$Census_Year-immigrant_table.df$Year_of_Immigration
immigrant_table_new.df=immigrant_table_new.df[,-c(1,12)]

immigrant_table_new.df=immigrant_table_new.df[,c(1:32,34,33)]


immigrant_table_new.df$Person_Weight
immigrant_table_new.df.expanded <- immigrant_table_new.df.expanded[rep(row.names(immigrant_table_new.df), immigrant_table_new.df$Person_Weight),]


immigrant_table_new.df.expanded =immigrant_table_new.df.expanded [complete.cases(immigrant_table_new.df.expanded ), ]#no NA



df2=immigrant_table_new.df.expanded
train<-sample(1:nrow(df2),nrow(df2)/2,rep=F)
train.df<-df2[train,]
test.df<-df2[-train,]


forestFit <- randomForest(as.factor(Sex) ~ .,
                      data=train.df, 
                      importance=TRUE, 
                      ntree=200)


pred2 <- predict(forestFit, test.df)

test.df$preds2<-pred2


with(test.df,table(Sex,preds2))
with(test.df,mean(preds2!=Sex,na.rm=T))

varImpPlot(forestFit)

```


##PCA

```{r}

df=immigrant_table_new.df.expanded[,-28]
mod.pc<-prcomp(df)
autoplot(prcomp(df))

autoplot(prcomp(df), data = immigrant_table_new.df.expanded, colour = 'Birthplace_final',
         loadings = TRUE, loadings.colour = 'blue',
         loadings.label = TRUE, loadings.label.size = 3)
```


##Tree
```{r}
df2=immigrant_table_new.df.expanded
train<-sample(1:nrow(df2),nrow(df2)/2,rep=F)
train.df<-df2[train,]
test.df<-df2[-train,]

model1 <- ctree(Birthplace_final ~ . , data = train.df)
plot(model1)
```









##knn.reg
```{r}

##knn_reg_function

knn_reg_model<-function(a){
  model.knn<-knn.reg(train.df,test.df,classes,k=a)
  test.df$preds<-sapply(model.knn[4],as.numeric)
  mse=with(test.df, mean((Dominance_Christan - preds)^2))
  mse
}

kval=seq(1,80,by=1)
i=1
mse_c=c()
while(i<=length(kval)){
  train.df <- immigrant_table_new.df[train,]
  test.df <- immigrant_table_new.df[-train,]
  classes<-with(train.df,Dominance_Christan)
  mse_c=c(mse_c,knn_reg_model(kval[i]))
  i=i+1
}
mse_c



test.df1=data.frame(flexibility=kval,mse=mse_c)



ggplot(test.df1,aes(flexibility,mse))+
    geom_point()+geom_smooth(aes(flexibility,mse))

kval[which.min(mse_c)]##optimal k-value

```

##Cross-Validation for KNN Regression

```{r}
kval=26

sampleSize<-nrow(immigrant_table_new.df.expanded)

mseCV <- function(data.df,kval,kfolds=10){
  folds <- sample(1:kfolds,sampleSize,rep=T)
  mse <- rep(0,kfolds)
  for(k in 1:kfolds){
    train.df1 <- data.df[folds !=k,]
    test.df1 <- data.df[folds==k,]
    classes<-with(train.df1,Christan_Percent)
    model.knn<-knn.reg(train.df1,test.df1,classes,k=kval)
    test.df1$preds<-sapply(model.knn[4],as.numeric)
    mse[k] <- with(test.df1,mean((Christan_Percent-preds)^2))
  }
  mean(mse)
}

#Results from Cross-Validation 
mse_knn_reg=mseCV(immigrant_table_new.df.expanded)
mse_knn_reg
```


#Classification


```{r}

set.seed(1234)
train<-sample(1:nrow(immigrant_table_new.df.expanded),nrow(immigrant_table_new.df.expanded )/2,rep=F)
train.df<-immigrant_table_new.df.expanded [train,]
test.df<-immigrant_table_new.df.expanded [-train,]
classes<-with(train.df,Christan_Percent )

kval<-2
model.knn<-knn(train.df,test.df,classes,k=kval,prob=T)
test.df$preds<-model.knn
with(test.df,mean(preds!=Dominance_Christan))##error rate 

##knn_function 

knn_model<-function(a){
train.df<-immigrant_table_new.df[train,]
test.df<-immigrant_table_new.df[-train,]
  model.knn<-knn(train.df,test.df,classes,k=a)
  test.df$preds<-model.knn 
  error_rate=with(test.df,mean(preds!=Dominance_Christan))
 error_rate
}
```


##KNN Cross Validation
```{r}
kval=2

mseCV <- function(data.df,kfolds=10){
  folds <- sample(1:kfolds,nrow(data.df),rep=T)
  mse <- rep(0,kfolds)
  for(k in 1:kfolds){
    train.df <- data.df[folds !=k,]
    test.df <- data.df[folds==k,]
    classes<-with(train.df,Dominance_Christan)
    model.knn <- knn(train.df,test.df,classes,k=kval,prob=T)
    test.df$preds<-model.knn
    err <- with(test.df, mean(preds!=Dominance_Christan))
    mse=c(mse,err)
  }
  mean(mse)
}


error_rate_knn=mseCV(immigrant_table_new.df)
error_rate_knn
```



##k mean clustering


```{r}
immigrant_table_new.df.expanded_new=mutate(immigrant_table_new.df.expanded,ifelse(immigrant_table_new.df.expanded$Christan_Percent>=mean(immigrant_table_new.df.expanded$Christan_Percent),"predominately_christian","not_predominately_christian"))

immigrant_table_new.df.expanded_new=immigrant_table_new.df.expanded_new[,-33]
immigrant_table_new.df.expanded_new$`ifelse(immigrant_table_new.df.expande...`=as.numeric(immigrant_table_new.df.expanded_new$`ifelse(immigrant_table_new.df.expande...`)##2=predominately chiristan 1=not predominately 
names(immigrant_table_new.df.expanded_new)[33]=c("Christian")

```

```{r}
pr.out=prcomp(immigrant_table_new.df.expanded_new, scale=TRUE)
biplot(pr.out, scale=0)
```
```{r}
pr.var=pr.out$sdev ^2
pve=pr.var/sum(pr.var)
which.max(pve)
pve[which.max(pve)]
names(pve)=names(immigrant_table_new.df.expanded_new)
pve
pve[order(pve,decreasing = T)]
```

```{r}
plot(pve, xlab="Principal Component", ylab="Proportion of Variance Explained ", ylim=c(0,1),type="b")
```

```{r}
small_data=immigrant_table_new.df.expanded_new
train<-sample(1:nrow(small_data),nrow(small_data)/20,rep=F)
train.df1<-small_data[train,]
test.df1<-small_data[-train,]

sd.data=scale(small_data)
hc.out=hclust(dist(sd.data))
km.out=kmeans(small_data,5,nstart=20)
km.clusters=km.out$cluster

cl1=small_data[which(km.clusters==1),]
cl2=small_data[which(km.clusters==2),]
cl3=small_data[which(km.clusters==3),]
cl4=small_data[which(km.clusters==4),]
cl5=small_data[which(km.clusters==5),]


summary(cl1)
summary(cl2)
summary(cl3)
summary(cl4)
summary(cl5)


hc.clusters=cutree(hc.out,5) 
table(hc.clusters)
par(mfrow=c(1,1))
plot(hc.out)
abline(h=139, col="red")



km.out=kmeans(train.df1,5,nstart=20)
km.clusters=km.out$cluster


table(km.clusters ,hc.clusters)#surprise?

```

```{r}
pr.out=prcomp(train.df1, scale=TRUE)

hc.out=hclust(dist(pr.out$x[,1:5]))
plot(hc.out, main="Hier. Clust. on First
Five Score Vectors ")
table(cutree(hc.out,5))

summary(pr.out)
```


```{r}
plot(train.df1, col=(km.out$cluster +1), main="K-Means Clustering Results with K=3", pch=20, cex=2)
ggplot(immigrant_table_new.df.expanded_new)
require(corrplot)
corrplot(cor(train.df1[,c(33,1:32)]),order = 'hclust',addrect = 5)
```

```{r}
hc.average=hclust(dist(train.df1), method="average")
```



```{r}
sd.data=scale(train.df1)


par(mfrow=c(1,3))
data.dist=dist(sd.data)

a=hclust(data.dist)
plot(hclust(data.dist), labels=nci.labs, main="Complete
Linkage", xlab="", sub="",ylab="")

plot(hclust(data.dist, method="average"), labels=nci.labs,
main="Average Linkage", xlab="", sub="",ylab="")


> plot(hclust(data.dist, method="single"), labels=nci.labs,
main="Single Linkage", xlab="", sub="",ylab="")

 hc.out=hclust(dist(pr.out$x[,1:5]))

```

