---
title: "Employment Response"
author: "Pony Ameri"
date: "4/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)

library(car)
library(plyr)
library(ipumsr)
library(labelled)
library(ggplot2)
library(mosaic)

usa_ddi <- read_ipums_ddi("/home/pelayo1/ImmigrationDUR/American Dream/usa_00008.xml")
usa_data <- read_ipums_micro(usa_ddi, verbose = FALSE)
usa_data <- usa_data %>%
  mutate(METRO = droplevels(as_factor(METRO)), BIRTHPLACE_factor= droplevels(as_factor(BPL)), LABFORCE1= droplevels(as_factor(LABFORCE)), SEX= droplevels(as_factor(SEX)),FOODSTMP= droplevels(as_factor(FOODSTMP)), RELATE= droplevels(as_factor(RELATE)),RELATED= droplevels(as_factor(RELATED)), MARST= droplevels(as_factor(MARST)), BPL = droplevels(as_factor(BPL)),BPLD = droplevels(as_factor(BPLD)),RACESINGD=droplevels(as_factor(RACESINGD)),HCOVANY = droplevels(as_factor(HCOVANY)),HCOPRIV = droplevels(as_factor(HCOVPRIV)), EDUCATION = droplevels(as_factor(GRADEATT)),EDUCATIOND= droplevels(as_factor(GRADEATTD)),EMPLOYMENT = droplevels(as_factor(EMPSTAT)), EMPLOYMENTD = droplevels(as_factor(EMPSTATD)), LABFORCE = droplevels(as_factor(LABFORCE)), IND = droplevels(as_factor(IND)), CLASSWKR = droplevels(as_factor(CLASSWKR)), CLASSWKRD = droplevels(as_factor(CLASSWKRD)), POVERTY = droplevels(as_factor(POVERTY)), MIGRATE1 = droplevels(as_factor(MIGRATE1)), MIGRATE1D = droplevels(as_factor(MIGRATE1D)), MIGPLAC1 = droplevels(as_factor(MIGPLAC1)), TRANWORK = droplevels(as_factor(TRANWORK)), OWNERSHP  = droplevels(as_factor(OWNERSHP)), HCOVPRIV1 = droplevels(as_factor(HCOVPRIV)))
usa_data$BPLD <-recode_factor(usa_data$BPLD, 'Egypt/United Arab Rep.' = "Egypt/UAE", 'Yemen Arab Republic (North)' = "Yemen")
usa_data$EMPLOYMENT <-recode_factor(usa_data$EMPLOYMENT, 'N/A' = "Armed Forces")
usa_data$LABFORCE1<-recode_factor(usa_data$LABFORCE1,'Yes, in the labor force'= "Yes",'No, not in the labor force'="No") 
usa_data$METRO<-recode_factor(usa_data$METRO,'Central / Principal city status unknown' = " Metro area, central", 'In metro area, central / principal city' = " Metro area, central",'In metro area, outside central / principal city' ="Metro area, outside central")
usa_data$OWNERSHP <-recode_factor(usa_data$OWNERSHP, 'Owned or being bought (loan)' = "Own", 'Rented' = "Rent")
usa_data <- usa_data %>% filter(YRIMMIG >1995,AGE >= 25)
usa_data$YRSINUSA <- usa_data$YEAR - usa_data$YRIMMIG
usa_data <- usa_data[usa_data$OWNERSHP != "0",]
usa_data$POVERTY <- as.numeric(usa_data$POVERTY)
usa_data <- usa_data[!is.na(usa_data$HCOVANY),]
usa_data <- usa_data %>% mutate(YRGROUP = ifelse(YRSINUSA <= 5,"0-5",
                                                 ifelse(YRSINUSA >5 & YRSINUSA <=10, "6-10",
                                                        ifelse(YRSINUSA >10 & YRSINUSA <=15, "11-15",
                                                               ifelse(YRSINUSA >15, ">15", "N/A")))),
                                Employment = ifelse(EMPLOYMENT == "Employed", 1,0), 
                                Laborforce = ifelse(LABFORCE1=="Yes",1,0),
                                Residence = ifelse(METRO == "Metro area, outside central",1,0), 
                                Ownhome = ifelse(OWNERSHP=="Own",1,0),
                                Poverty = ifelse(POVERTY >150, 1,0),
                                Health = ifelse(HCOVPRIV1 == "With private health insurance coverage", 1, 0),
                                Foodstamp = ifelse(FOODSTMP=="No",1,0),
                                Race = ifelse(RACESING==1, "White", "Non-white"),
                                Citizenship = ifelse(CITIZEN == c(1,2), "Citizen", "Non-citizen"))
usa_data1 <- usa_data[!is.na(usa_data$Race),]

```

```{r cars}
                    

mosaicplot(Employment~Race, data=usa_data)
mosaicplot(Employment~SEX, data=usa_data)
mosaicplot(Employment~Citizenship, data=usa_data)

ggplot(usa_dat, aes(YRSINUSA))
```


```{r cars}
dream<- glm(Employment~ AGE + Race +SEX + YRSINUSA+Citizenship +MIGRATE1+BPLD +Residence +FAMSIZE + MARST+ GRADEATT, family=binomial(link="logit"), data=usa_data1)
summary(dream)
exp(dream$coefficients)
confint(dream)
1-pchisq(dream$deviance,dream$df.residual) #lack of significant fit

dream1<- glm(Employment~ AGE + Race +SEX + YRSINUSA+Citizenship +MIGRATE1+BPLD +Residence +FAMSIZE + MARST+ GRADEATT, family=quasibinomial(link="logit"), data=usa_data1)
summary(dream1)

```

Race,Migration within state,birthplace Iran, sudan, Somalia , maristal status married or separated are not significant predictors

The average number of people employed increases by 8% for every additional year living in the United states, after controlling for the effect of age, race, citizenship, birthplace, Migration status, residence, grade attained and marital status. 

```{r}

```
Statistically significant evidence of lack of fit (p<.001).

