---
title: "IRT"
output: html_document
---

IRT
```{r}
library(readr)
library(dplyr)

library(car)
library(plyr)
library(ipumsr)
library(labelled)
library(ggplot2)
library(mosaic)


usa_ddi <- read_ipums_ddi("/home/pelayo1/ImmigrationDUR/American Dream/usa_00008.xml")
usa_data <- read_ipums_micro(usa_ddi, verbose = FALSE)
usa_data <- usa_data %>%
  mutate(METRO = droplevels(as_factor(METRO)), BIRTHPLACE_factor= droplevels(as_factor(BPL)), LABFORCE1= droplevels(as_factor(LABFORCE)), SEX= droplevels(as_factor(SEX)),FOODSTMP= droplevels(as_factor(FOODSTMP)), RELATE= droplevels(as_factor(RELATE)),RELATED= droplevels(as_factor(RELATED)), MARST= droplevels(as_factor(MARST)), BPL = droplevels(as_factor(BPL)),BPLD = droplevels(as_factor(BPLD)), CITIZEN = droplevels(as_factor(CITIZEN)),RACESING=droplevels.factor(RACESING),RACESINGD=droplevels(as_factor(RACESINGD)),HCOVANY = droplevels(as_factor(HCOVANY)),HCOPRIV = droplevels(as_factor(HCOVPRIV)), EDUCATION = droplevels(as_factor(GRADEATT)),EDUCATIOND= droplevels(as_factor(GRADEATTD)),EMPLOYMENT = droplevels(as_factor(EMPSTAT)), EMPLOYMENTD = droplevels(as_factor(EMPSTATD)), LABFORCE = droplevels(as_factor(LABFORCE)), IND = droplevels(as_factor(IND)),
 CLASSWKR = droplevels(as_factor(CLASSWKR)), CLASSWKRD = droplevels(as_factor(CLASSWKRD)), POVERTY = droplevels(as_factor(POVERTY)), MIGRATE1 = droplevels(as_factor(MIGRATE1)), MIGRATE1D = droplevels(as_factor(MIGRATE1D)), MIGPLAC1 = droplevels(as_factor(MIGPLAC1)), TRANWORK = droplevels(as_factor(TRANWORK)), OWNERSHP  = droplevels(as_factor(OWNERSHP)), HCOVPRIV1 = droplevels(as_factor(HCOVPRIV)), RACE= droplevels(as_factor(RACESING)))
usa_data$BPLD <-recode_factor(usa_data$BPLD, 'Egypt/United Arab Rep.' = "Egypt/UAE", 'Yemen Arab Republic (North)' = "Yemen")
usa_data$EMPLOYMENT <-recode_factor(usa_data$EMPLOYMENT, 'N/A' = "Armed Forces")
usa_data$LABFORCE1<-recode_factor(usa_data$LABFORCE1,'Yes, in the labor force'= "Yes",'No, not in the labor force'="No") 
usa_data$METRO<-recode_factor(usa_data$METRO,'Central / Principal city status unknown' = " Metro area, central", 'In metro area, central / principal city' = " Metro area, central",'In metro area, outside central / principal city' ="Metro area, outside central")
usa_data$OWNERSHP <-recode_factor(usa_data$OWNERSHP, 'Owned or being bought (loan)' = "Own", 'Rented' = "Rent")
usa_data <- usa_data %>% filter(YRIMMIG >1995,AGE >= 25)
usa_data$YRSINUSA <- usa_data$YEAR - usa_data$YRIMMIG
usa_data <- usa_data[usa_data$OWNERSHP != "0",]
usa_data$POVERTY <- as.numeric(usa_data$POVERTY)
usa_data <- usa_data[!is.na(usa_data$HCOVANY),]
usa_data <- usa_data %>% mutate(YRGROUP = ifelse(YRSINUSA <= 5,"0-5",
                                    ifelse(YRSINUSA >5 & YRSINUSA <=10, "6-10",
                                   ifelse(YRSINUSA >10 & YRSINUSA <=15, "11-15",
                                    ifelse(YRSINUSA >15, ">15", "N/A")))),
                                Employment = ifelse(EMPLOYMENT == "Employed", 1,0), 
                              Laborforce = ifelse(LABFORCE1=="Yes",1,0),
                              Residence = ifelse(METRO == "Metro area, outside central",1,0), 
                              Ownhome = ifelse(OWNERSHP=="Own",1,0),
                              Poverty = ifelse(POVERTY >150, 1,0),
                              Health = ifelse(HCOVPRIV1 == "With private health insurance coverage", 1, 0),
                              Foodstamp = ifelse(FOODSTMP=="No",1,0),
                              White = ifelse(RACE==1,1,0),
                              Citizen = ifelse(CITIZEN == ' Not a citizen', 0, 1))

```
                 
                 
                 
```{r}
library(readr)
library(dplyr)
library(car)
library(plyr)
library(ipumsr)
library(labelled)
library(ggplot2)
library(mosaic)

usa_ddi1 <- read_ipums_ddi("/home/pelayo1/ImmigrationDUR/American Dream/usa_00013.xml")
usadata <- read_ipums_micro(usa_ddi1, verbose = FALSE)
usadata <-usadata %>% filter(AGE > 25)
usadata <- na.omit(usadata)

usadata <- usadata %>%
  mutate(METRO = droplevels(as_factor(METRO)), BIRTHPLACE_factor= droplevels(as_factor(BPL)), LABFORCE1= droplevels(as_factor(LABFORCE)), SEX= droplevels(as_factor(SEX)),FOODSTMP= droplevels(as_factor(FOODSTMP)), RELATE= droplevels(as_factor(RELATE)),RELATED= droplevels(as_factor(RELATED)), MARST= droplevels(as_factor(MARST)), BPL = droplevels(as_factor(BPL)),BPLD = droplevels(as_factor(BPLD)), CITIZEN = droplevels(as_factor(CITIZEN)),RACESING=droplevels.factor(RACESING),RACESINGD=droplevels(as_factor(RACESINGD)),HCOVANY = droplevels(as_factor(HCOVANY)),HCOPRIV = droplevels(as_factor(HCOVPRIV)), EDUCATION = droplevels(as_factor(GRADEATT)),EDUCATIOND= droplevels(as_factor(GRADEATTD)),EMPLOYMENT = droplevels(as_factor(EMPSTAT)), EMPLOYMENTD = droplevels(as_factor(EMPSTATD)), LABFORCE = droplevels(as_factor(LABFORCE)), IND = droplevels(as_factor(IND)),
 CLASSWKR = droplevels(as_factor(CLASSWKR)), CLASSWKRD = droplevels(as_factor(CLASSWKRD)), MIGRATE1 = droplevels(as_factor(MIGRATE1)), MIGRATE1D = droplevels(as_factor(MIGRATE1D)), MIGPLAC1 = droplevels(as_factor(MIGPLAC1)), TRANWORK = droplevels(as_factor(TRANWORK)), OWNERSHP  = droplevels(as_factor(OWNERSHP)), HCOVPRIV1 = droplevels(as_factor(HCOVPRIV)), RACE= droplevels(as_factor(RACESING)))

usadata$POVERTY = as.numeric(usadata$POVERTY)
usadata <- usadata[!is.na(usadata$HCOVPRIV1),]
usadata <- usadata[usadata$OWNERSHP != "N/A",]
usadata$LABFORCE1<-recode_factor(usadata$LABFORCE1,'Yes, in the labor force'= "Yes",'No, not in the labor force'="No") 
usadata$OWNERSHP <-recode_factor(usadata$OWNERSHP, 'Owned or being bought (loan)' = "Own", 'Rented' = "Rent")
usadata$YRSINUSA <- usadata$YEAR - usadata$YRIMMIG
usadata <- usadata %>% mutate(YRGROUP = ifelse(YRSINUSA <= 5,"0-5",
                                    ifelse(YRSINUSA >5 & YRSINUSA <=10, "6-10",
                                   ifelse(YRSINUSA >10 & YRSINUSA <=15, "11-15",
                                    ifelse(YRSINUSA >15, ">15", "N/A")))),
                              Employment1 = ifelse(EMPLOYMENT == "Employed", 1,0), 
                              Laborforce1 = ifelse(LABFORCE1=="Yes",1,0),
                              Residence1 = ifelse(METRO == "Metro area, outside central",1,0), 
                              Ownhome1 = ifelse(OWNERSHP=="Own",1,0),
                              Poverty1 = ifelse(POVERTY > 150, 1,0),
                              Health1 = ifelse(HCOVPRIV1 == "With private health insurance coverage", 1, 0),
                              Foodstamp1 = ifelse(FOODSTMP=="No",1,0),
                              White1 = ifelse(RACE==1,1,0),
                              Citizen1 = ifelse(CITIZEN == ' Not a citizen', 0, 1))


```



```{r}
######### IRT ########
View(usadata)
#### Immigrants
par(mfrow=c(2,2))
immigus <- data.frame(Emp= rep(usa_data$Employment, usa_data$PERWT),
                 Pov = rep(usa_data$Poverty, usa_data$PERWT),
                 Fs = rep(usa_data$Foodstamp, usa_data$PERWT),
                Home = rep(usa_data$Ownhome,usa_data$PERWT),
                Hlth = rep(usa_data$Health, usa_data$PERWT),
                Years = rep(usa_data$YRGROUP,usa_data$PERWT))
View(immigus)
library(ltm)
library(msm)
descript(immigus)
ADmodel <- ltm(immigus[,1:5]~z1, IRT.param = T)
plot(ADmodel, main = "Immigrant Success Score", type="ICC")
plot(ADmodel, type="IIC",main = "Immigrant Item Information Curve")


##### Non-Immigrants
library(data.table)
set.seed(57)
samp <- data.table(usadata)
nonimm<-samp[sample(nrow(samp), 10000)]
nonimm.subset <- data.frame(Emp= rep(nonimm$Employment1, nonimm$PERWT),
                 Pov = rep(nonimm$Poverty1, nonimm$PERWT),
                 Fs = rep(nonimm$Foodstamp1, nonimm$PERWT),
                Home = rep(nonimm$Ownhome1,nonimm$PERWT),
                Hlth = rep(nonimm$Health1, nonimm$PERWT))

library(ltm)
library(msm)
descript(nonimm.subset)
ADmodel.a <- ltm(nonimm.subset~z1, IRT.param = T)
plot(ADmodel.a, main = "Non- immigrant Success Score", type="ICC")
plot(ADmodel.a, type="IIC",main = "Immigrant Item Information Curve")

```

The Item characteristic curve illustrates that owning a home, being above the poverty level and not being recipients of food stamps are in general very difficult to attain. 

A high discrimination parameter value suggests an item that has a high ability to differentiate subjects.
In practice, a high discrimination parameter value means that the probability of a correct response increases more rapidly as the ability (latent trait) increases.

The discrimination parameter for the employment, Poverty, Food stamp recipiency, Owning a home and having private health insurance is 0.9, 2.89, 1.86, 0.92 and 2.11 respectively.

This means that the probability of an immigrant from these countries to be above the poverty level, not require assistance purchasing food, and have private health insurance coverage increases more rapidly as the ability increases, compared to the probability of being employed and owning a home.

```{r}
#### Success score over time for immigrants
par(mfrow = c(2,2))
immigus1 <- immigus %>% filter(Years=="0-5")
AD <- ltm(immigus1[,1:5]~z1, IRT.param = T)
plot(AD, main = "Immigrant Success Score for first 5 years", type="ICC")
plot(AD, type="IIC",main = "Immigrant Item Information Curve for first 5 years")

immigus2 <- immigus %>% filter(Years=="6-10")
AD1 <- ltm(immigus2[,1:5]~z1, IRT.param = T)
plot(AD1, main = "Immigrant Success Score from 6-10 years", type="ICC")
plot(AD1, type="IIC",main = "Immigrant Item Information Curve from 6-10 years")

immigus3 <- immigus %>% filter(Years=="11-15")
AD2 <- ltm(immigus3[,1:5]~z1, IRT.param = T)
plot(AD2, main = "Immigrant Success Score from 11-15 years", type="ICC")
plot(AD2, type="IIC",main = "Immigrant Item Information Curve from 11-15 years")

immigus4 <- immigus %>% filter(Years==">15")
AD3 <- ltm(immigus4[,1:5]~z1, IRT.param = T)
plot(AD3, main = "Immigrant Success Score for over 15 years", type="ICC")
plot(AD3, type="IIC",main = "Immigrant Item Information Curve for over 15 years")
```


```{r}

par(mfrow = c(2,2))
us1 <-usa_data%>% filter(BPLD == "Yemen")

usYemen <- data.frame(Emp= rep(us1$Employment, us1$PERWT),
                 Pov = rep(us1$Poverty, us1$PERWT),
                 Fs = rep(us1$Foodstamp, us1$PERWT),
                Home = rep(us1$Ownhome,us1$PERWT),
                Hlth = rep(us1$Health, us1$PERWT))

descript(usYemen)
ADmodel1 <- ltm(usYemen~z1, IRT.param = T)
summary(ADmodel1)
plot(ADmodel1, main = "Success Score for Yemen", type="ICC")
plot(ADmodel1, type="IIC",main = "Yemen Item Information Curve")

#Iran
us2 <-usa_data%>% filter(BPLD == "Iran")

usIran<- data.frame(Emp= rep(us2$Employment, us2$PERWT),
                 Pov = rep(us2$Poverty, us2$PERWT),
                 Fs = rep(us2$Foodstamp, us2$PERWT),
                Home = rep(us2$Ownhome,us2$PERWT),
                Hlth = rep(us2$Health, us2$PERWT))

descript(usIran)
ADmodel2 <- ltm(usIran~z1, IRT.param = T)
summary(ADmodel2)
plot(ADmodel2, main = "Success Score for Iran", type="ICC")
plot(ADmodel2, type="IIC",main = "Iran Item Information Curve")

#Egypt/UAE
us3 <-usa_data%>% filter(BPLD == "Egypt/UAE")

usEgypt<- data.frame(Emp= rep(us3$Employment, us3$PERWT),
                 Pov = rep(us3$Poverty,us3$PERWT),
                 Fs = rep(us3$Foodstamp,us3$PERWT),
                Home = rep(us3$Ownhome,us3$PERWT),
                Hlth = rep(us3$Health, us3$PERWT))

descript(usEgypt)
ADmodel3 <- ltm(usEgypt~z1, IRT.param = T)
summary(ADmodel3)
plot(ADmodel3, main = "Success Score for Egypt", type="ICC")
plot(ADmodel3, type="IIC",main = "Egypt Item Information Curve")


#Iraq
us4 <-usa_data%>% filter(BPLD == "Iraq")

usIraq<- data.frame(Emp= rep(us4$Employment, us4$PERWT),
                 Pov = rep(us4$Poverty,us4$PERWT),
                 Fs = rep(us4$Foodstamp, us4$PERWT),
                Home = rep(us4$Ownhome,us4$PERWT),
                Hlth = rep(us4$Health, us4$PERWT))

descript(usIraq)
ADmodel4 <- ltm(usIraq~z1, IRT.param = T)
summary(ADmodel4)
plot(ADmodel4, main = "Success Score for Iraq", type="ICC")
plot(ADmodel4, type="IIC",main = "Iraq Item Information Curve")

#Somalia
us5 <-usa_data%>% filter(BPLD == "Somalia")

usSomalia<- data.frame(Emp= rep(us5 $Employment, us5 $PERWT),
                 Pov = rep(us5 $Poverty,us5 $PERWT),
                 Fs = rep(us5 $Foodstamp,us5 $PERWT),
                Home = rep(us5 $Ownhome,us5 $PERWT),
                Hlth = rep(us5 $Health, us5 $PERWT))

descript(usSomalia)
ADmodel5 <- ltm(usSomalia~z1, IRT.param = T)
summary(ADmodel5)
plot(ADmodel5, main = "Success Score for Somalia", type="ICC")
plot(ADmodel5, type="IIC",main = "Somalia Item Information Curve")

#Syria
us6 <-usa_data%>% filter(BPLD == "Syria")

usSyria<- data.frame(Emp= rep(us6$Employment, us6$PERWT),
                 Pov = rep(us6$Poverty,us6$PERWT),
                 Fs = rep(us6$Foodstamp,us6$PERWT),
                Home = rep(us6$Ownhome,us6$PERWT),
                Hlth = rep(us6$Health,us6$PERWT))

descript(usSyria)
ADmodel6 <- ltm(usSyria~z1, IRT.param = T)
summary(ADmodel6)
plot(ADmodel6, main = "Success Score for Syria", type="ICC")
plot(ADmodel6, type="IIC",main = "Syria Item Information Curve")

#Libya
us7 <-usa_data%>% filter(BPLD == "Libya")

usLibya<- data.frame(Emp= rep(us7$Employment, us7$PERWT),
                 Pov = rep(us7$Poverty,us7$PERWT),
                 Fs = rep(us7$Foodstamp, us7$PERWT),
                Home = rep(us7$Ownhome,us7$PERWT),
                Hlth = rep(us7$Health,us7$PERWT))

descript(usLibya)
ADmodel7 <- ltm(usLibya~z1, IRT.param = T)
summary(ADmodel7)
plot(ADmodel7, main = "Success Score for Libya", type="ICC")
plot(ADmodel7, type="IIC",main = "Libya Item Information Curve")

#Sudan
us8 <-usa_data%>% filter(BPLD == "Sudan")

usSudan<- data.frame(Emp= rep(us8 $Employment, us8 $PERWT),
                 Pov = rep(us8 $Poverty,us8 $PERWT),
                 Fs = rep(us8 $Foodstamp, us8 $PERWT),
                Home = rep(us8 $Ownhome,us8 $PERWT),
                Hlth = rep(us8 $Health, us8 $PERWT))

descript(usSudan)
ADmodel8 <- ltm(usSudan~z1, IRT.param = T)
summary(ADmodel8)
plot(ADmodel8, main = "Success Score for Sudan", type="ICC")
plot(ADmodel8, type="IIC",main = "Sudan Item Information Curve")

table(usa_data$BPLD)
```


successscore - dependent variable 

compare from country to country 
year
how do success scores differ for the countries in question and the US


```{r}
# z-score extraction
Dream1<-factor.scores(ADmodel)$score.dat
Dream1$Employment <-Dream1$Emp
Dream1$Poverty <-Dream1$Pov
Dream1$Foodstamp <-Dream1$Fs
Dream1$Ownhome <-Dream1$Home
Dream1$Health <-Dream1$Hlth
use<-merge(usa_data, Dream1, by= c("Employment","Poverty","Foodstamp","Ownhome","Health")) #new dataset

View(usa_data)
dats<-use$success <-use$z1

# modelling

library(leaps)

bestss <- regsubsets(success ~ FAMSIZE + SEX + AGE + MARST + CITIZEN + YRIMMIG + HCOVANY + LABFORCE +INCTOT  + MIGRATE1 + YRSINUSA, data = use, method = "exhaustive", nbest = 3)
summary.bestss <- summary(bestss, matrix.logical = TRUE) # matrix.logical term makes table of included terms more readable
summary.bestss
coef(bestss, 1:15)

plot(bestss, scale = "adjr2", main = "Adjusted R^2")
plot(bestss, scale = "Cp", main = "Mallow's Cp")

# Which model has best adjusted R^2
which.max(summary.bestss$adjr2)
summary.bestss$which[7,]

# Which model has best Mallow's Cp
which.min(summary.bestss$cp)
summary.bestss$which[7,]

```



```{r}
# z-score extraction
Dream2<-factor.scores(ADmodel1)$score.dat
Dream2$Employment1 <-Dream2$Emp
Dream2$Poverty1 <-Dream2$Pov
Dream2$Foodstamp1 <-Dream2$Fs
Dream2$Ownhome1 <-Dream2$Home
Dream2$Health1 <-Dream2$Hlth

use1<-merge(nonimm, Dream2, by= c("Employment1","Poverty1","Foodstamp1","Ownhome1","Health1")) #new dataset
use1$success <-use1$z1

# modelling

bestss1 <- regsubsets(success ~ FAMSIZE + SEX + AGE + MARST + HCOVANY + LABFORCE +INCTOT  + MIGRATE1, data = use1, method = "exhaustive", nbest = 3)
summary.bestss1 <- summary(bestss1, matrix.logical = TRUE) # matrix.logical term makes table of included terms more readable
summary.bestss1
coef(bestss1, 1:15)

plot(bestss1, scale = "adjr2", main = "Adjusted R^2")
plot(bestss1, scale = "Cp", main = "Mallow's Cp")

# Which model has best adjusted R^2
which.max(summary.bestss1$adjr2)
summary.bestss1$which[7,]

# Which model has best Mallow's Cp
which.min(summary.bestss1$cp)
summary.bestss1$which[7,]

```


